The Rockerverse: Images, Packages, and Applications for Containerization with R

### 2020-01-27 manuscript.Rmd
wrap up deployment/CD section, closes #36

**\CRANpkg{stevedore}** ([https://cran.r-project.org/package=stevedore](https://cran.r-project.org/package=stevedore)) is currently the most powerful Docker client in R.
It interfaces with the Docker daemon over the Docker HTTP API\footnote{\href{https://docs.docker.com/engine/api/latest/}{https://docs.docker.com/engine/api/latest/}} via a Unix socket on Linux or MacOS, over a named pipe on Windows, or over an HTTP/TCP connection.
The package is the only one not using system calls to the `docker` CLI for managing images and containers and easily exposes connections to remote Docker daemons, which has to be configured on the Docker level otherwise.
Using the API gives access to more information and is system independent and likely more reliable than parsing command line output.
\CRANpkg{stevedore}'s own interface is automatically generated based on the OpenAPI specification of the Docker daemon, but still similar to the Docker CLI.
The interface is similar to R6 objects, in that a `stevedore_object` has a number of functions attached to it that can be called, and multiple specific versions of the Docker API can be supported thanks to the automatic generation\footnote{See \href{https://github.com/richfitz/stevedore/tree/master/inst/spec}{https://github.com/richfitz/stevedore/tree/master/inst/spec} and \href{https://github.com/richfitz/stevedore/blob/master/development.md}{https://github.com/richfitz/stevedore/blob/master/development.md}.}.

**\CRANpkg{AzureContainers}** is an interface to a number of container-related services in Microsoft's [Azure Cloud](https://azure.microsoft.com/), see Section&nbsp;\nameref{azurecontainers}.
While it is mainly intended for working with Azure, as a convenience feature it includes lightweight, cross-platform shells to Docker and Kubernetes (tools `kubectl` and `helm`).
These can be used to create and manage arbitrary Docker images and containers, as well as Kubernetes clusters on any platform or cloud service.

@@

## Continuous integration, continuous delivery, and deployment
\label{cicd}
\label{deployment}

The cloud is the natural environment of containers, and subsequently containers are the go-to mechanism to deploy R server applications.
More and more continuous integration (CI) and continuous delivery (CD) services also use containers, opening up new options for use.
The controlled nature of containers, i.e., the possibility to define the software environment very well, even on remote machines, is crucial here, for example to match test or build environments with production environments.

As a first example, \CRANpkg{Shiny} is the most popular package for creating interactive online dashboards with R and it enables users with very diverse backgrounds to create stable and user friendly web applications.
The ShinyProxy ([https://www.shinyproxy.io/](https://www.shinyproxy.io/)) is an open source tool to deploy Shiny apps in an enterprise context.
They feature single sign-on, but also in scientific use cases \citep[e.g., ][]{savini_epiexplorer_2019,glouzon_structurexplor_2017}.
ShinyProxy uses Docker containers to isolate user sessions and to achieve scalability for multi-user scenarious with multiple apps.
ShinyProxy itself is written in Java to accomodate corporate requirements and may itself run in a container for stability and availability.
The tool is build on ContainerProxy ([https://www.containerproxy.io/](https://www.containerproxy.io/)), which provides similar features for executign long-running R jobs or interactive R sessions.
The started containers can run on a regular Docker host but also in clusters.

@@

- [dynwrap](https://github.com/dynverse/dynwrap_containers/blob/master/.travis.yml) [@rcannood]
  - For this project, we use travis-ci to build rocker-derived containers, test them, and only push them to Docker&nbsp;Hub (from travis-ci.org) if the integration tests succeed.

For example, the **[`golem`](https://github.com/ThinkR-open/golem)** package makes an heavy use of `dockerfiler` when it comes to creating the `Dockerfile` for building production-grade Shiny applications and deploying them.

- [`googleComputeEngineR`](https://cloudyr.github.io/googleComputeEngineR/) (function `gce_vm_template()`)
  - To enable quick deployments of key R services such as RStudio and Shiny onto cloud virtual machines (VMs), this package utilises Dockerfiles to move the labour of setting up those services from the user to a premade Docker image. For example, by specifying the template `template="rstudio"` in `gce_vm_template()/gce_vm()` an up to date RStudio Service image is launched.  Specifying `template="rstudio-gpu"` will launch an RStudio Server image with a GPU attached, etc.  
- [`analogsea`](https://github.com/sckott/analogsea) (digital ocean R client)
- [`plumber`](https://www.rplumber.io/docs/hosting.html#docker)
- [production deployment of neural networks](https://github.com/tmobile/r-tensorflow-api) [@jnolis, @nolistic]
- Kubernetes - A popular platform for managing Docker containers is [Kubernetes](https://kubernetes.io/), which is used for a wide variety of containerized applications. It may be your organisation has a Kubernetes cluster already for other applications. Docker containers are used within Kubernetes clusters to hold native code, for which Kubernetes creates a framework around network connections and scaling of resources up and down. An introduction on using [Kubernetes with R is published at this blog post](https://code.markedmondson.me/r-on-kubernetes-serverless-shiny-r-apis-and-scheduled-scripts/).

\label{azurecontainers}
**\CRANpkg{AzureContainers}** is an umbrella package providing interfaces three commercial services of Microsoft's Azure Cloud, namely [Container Instances](https://azure.microsoft.com/en-us/services/container-instances/) for running individual containers, [Container Registry](https://azure.microsoft.com/en-us/services/container-registry/) for private image distribution, and [Kubernetes Service](https://azure.microsoft.com/en-us/services/kubernetes-service/) for orchestrated deployments.
While a package like \CRANpkg{plumber} provides the infrastructure for turning an R workflow a service, for production purposes it is usually necessary to take into account scalability, reliability and ease of management.
A _cluster_ of containers, orchestrated as a single deployment, e.g., with [Kubernetes](https://kubernetes.io), can mitigate limitations on request volumes or a container occupied with computationally intensive task.
A cluster features load-balancing, autoscaling of containers across numerous servers (in the cloud or on premise), and restarting failed ones.
AzureContainers provides an R-based interface to these features for Azure.
On the client side, it provides simple wrappers to Docker and to Kubernetes' management tools `kubectl` and `helm`.
The package simplifies complex infrastructure management to a number of R function calls, given an Azure account with sufficient credit\footnote{See \emph{"Deploying a prediction service with Plumber"} vignette for details:  \href{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01_plumber_deploy.html}{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01\_plumber\_deploy.html}.}.

The prevelance of Docker in industry naturally leads to usage of containers with R in such settings as well, as customers already manage platforms in Docker containers.
These products often entail a high amount of open source software in combination with proprietary layers addign the relevant commercializable features.
One such example is RStudio's data science platform [RStudio Team](https://rstudio.com/products/team/).
It allows teams of data scientists and their respective IT/DevOps groups to develop and deploy code in R and Python around the RStudio Open Source Server inside of Docker images, without requiring users to learn new tools or directly interact with containers.
The best practices for [running RStudio with Docker containers](https://support.rstudio.com/hc/en-us/articles/360021594513-Running-RStudio-with-Docker-containers) as well as [Docker images](https://github.com/rstudio/rstudio-docker-products) for RStudio's commerical products are openly available.

@@

It deterministically builds a Docker image by parsing the contents of a repository, e.g., project dependency configurations or simple configuration files\footnote{See supported file types at \href{https://repo2docker.readthedocs.io/en/latest/config_files.html}{https://repo2docker.readthedocs.io/en/latest/config_files.html}.}.

@@

\citet{gentleman_statistical_2007} coined the term **Research Compendium** for a dynamic document together with supporting data and code.
They use the R package system \cite{core_writing_1999} as for the functional prototype to structuring, validation, and distribution of research compendia.
This concept has been taken up and extended\footnote{See full literature list at \href{https://research-compendium.science/}{https://research-compendium.science/}.}, not the least by applying containerisation and other methods for managing computing environments---see Section&nbsp;\nameref{envs}.
Containers give the researcher an isolated environment to assemble these research pipelines with specific versions of software.
Research workflows in containers are safe from contamination from other activities occuring on the researcher's computer, for example the installation of newest version of packages for teaching demonstrations.
Given the users in this scenario, i.e., often academics with limited formal software development training, templates and assistance with containers around research compendia is essential.
In many fields we see that a typical unit of research for a container is a research report or journal article, where the container holds the compendium, or self-contained set of data (or connections to data elsewhere) and code files needed to fully reproduce the article \citep{marwick_packaging_2018}.
The package \pkg{rrtools} ([https://github.com/benmarwick/rrtools](https://github.com/benmarwick/rrtools)) provides a template and convienence functions to apply good practices for research compendia, including a starter `Dockerfile`.
Images of compendium containers can be hosted on services such as Docker&nbsp;Hub for convienient sharing among collaborators and others. 
Similarly, packages such as \pkg{containerit} and \pkg{dockerfiler} can be used to manage the `Dockerfile` to be archived with a compendium on a data repository (e.g. [Zenodo](https://zenodo.org/), [Dataverse](https://dataverse.org/), [Figshare](https://figshare.com/), [OSF](https://osf.io/)).
A typial compendium's `Dockerfile` will pull a rocker image fixed to a specific version of R, and install R packages from the MRAN repository to ensure the package versions are tied to a specific date, rather than the most recent version.
A more extreme case is the _dynverse_ project ([https://dynverse.org/](https://dynverse.org/)), which packages over 50 computational methods with different environments (R, Python, C++, etc.) in Docker images, which can be executed from R.
Future researchers can download the compendium from the repository and run the included `Dockerfile` to build a new image that recreates the computational environment used to produce the original research results.
If building the image fails, the human-readable instructions in a `Dockerfile` are the starting point for rebuilding the environment.
Further safeguarding practices are currently under development or not part of common practice yet, such as preservation of images \citep{emsley_framework_2018} and storing them alongside `Dockerfile`s \citep[cf.][]{nust_opening_2017}, or pinning of system libraries beyond the tagged base images, which may be seen as stable or dynamic depending on the applied time scale \citep[see discussion on `debian:testing` base image in][]{RJ-2017-065}.

A recommendation of the recent National Academies' report on _Reproducibility and Replicability in Science_ is that journals _"consider ways to ensure computational reproducibility for publications that make claims based on computations"_ \citep{NASEM2019}.
In fields such as Political Science and Economics, journals are increasingly adopting policies requiring authors to publish the code and data required to reproduce computational findings reported in published manuscripts, subject to independent verification \citep{Jacoby2017,Vilhuber2019,Alvarez2018,Christian2018,Eubank2016,King1995}.
Problems with the computational environment, installation and availability of software dependencies are common.
R is gaining popularity in these communities, such as creating a research compendium.
In a sample of 105 replication packages published by the _American Journal of Political Science_ (AJPS) over 65% use R.
The NSF-funded Whole Tale project uses the Rocker Project community images with the goal of improving the reproducibility of published research artifacts and simplifying the publication and verification process for both authors and reviewers by reducing errors and time spent specifying the environment.

@@

Nonetheless, at least on the level of R packages some **consolidation** seems in order, e.g., to reduce the number of packages creating `Dockerfile`s from R code or controlling the Docker daemon with R code.
It remains to be seen which approach to control Docker, via the Docker API as \pkg{stevedore} or via system calls as \pkg{dockyard}/\pkg{docker}/\pkg{dockr}, is more sustainable, or if the question will be answered by the endurance of maintainers and sufficient funding.
Similarly, the capturing of environments and their serialization in form of a `Dockerfile` currently happens at different levels of abstraction and re-use of functionality seems reasonable, e.g., \pkg{liftr} could generate the environment with \pkg{containerit}, which in turn may use \pkg{dockerfiler} for low level R objects representing a `Dockerfile` and its instructions.
In this consolidation, the Rocker&nbsp;Project could play the role of coordinating entity.
Though for the moment, the sign of the times points to more experimentation and feature growth, e.g., images for GPU-based computing and artificial intelligence.
Even with coding being more and more accepted as a required, and achievable skill, an easier access, for example by exposing containerisation benefits via simple user interfaces in the users' IDE, could be an important next step.
Currently containerisation happens more in the background at the system level.
Furthermore, a maturing of the Rockerverse packages for managing containers may lead to their adoption where currently manual coding is required, e.g. in the case of \CRANpkg{RSelentium} or \CRANpkg{drake} (see Sections&nbsp;\nameref{rselenium} and \nameref{drake} respectively).

@@

The ordering of authors following DN and DE is alphabetical.
DN conceived the article idea, \href{https://github.com/nuest/rockerverse-paper/issues/3}{initialised the formation of the writing team}, wrote sections not mentioned below, and revised all sections.
DB wrote the section on \pkg{outsider}.
GD contributed the CARD.com use case.
DE wrote the introduction and the section about Containerization and Rocker and reviewed all sections.
RC \& EH contributed to the section on interfaces for Docker in R.
DC contributed content on Gigantum.
ME contributed to the section on processing and deployment to cloud services.
CF wrote paragraphs about \pkg{r-online}, \pkg{dockerfiler}, \pkg{r-ci} and \pkg{r-db}.
SL contributed content on RStudio's usage of Docker.
BM wrote the section on research compendia and made the project Binder-ready.
HN \& JN co-wrote the section on enterprise production environments (\pkg{googleComputeEngineR}, Google Cloud Run, Kubernetes).
KR wrote the section about \pkg{holepunch}.
NR wrote paragraphs about shared work environments and GPUs.
LS \& NT wrote the section on Bioconductor.
CW wrote the sections on Whole Tale and contributed the publication reproducibility audit use case.
NX contributed content on \pkg{liftr}.
All authors approved the final version.
This articles was collaboratively written at \href{https://github.com/nuest/rockerverse-paper/}{https://github.com/nuest/rockerverse-paper/}.
The \href{https://github.com/nuest/rockerverse-paper/graphs/contributors}{contributors page}, \href{https://github.com/nuest/rockerverse-paper/commits/master}{commit history}, and \href{https://github.com/nuest/rockerverse-paper/issues/}{discussion issues} provide a detailed view on the respective contributions.

>>>>

**\CRANpkg{stevedore}** ([https://cran.r-project.org/package=stevedore](https://cran.r-project.org/package=stevedore)) is currently the most powerful Docker client in R.
It interfaces with the Docker daemon over the Docker HTTP API\footnote{\href{https://docs.docker.com/engine/api/latest/}{https://docs.docker.com/engine/api/latest/}} via a Unix socket on Linux or MacOS, over a named pipe on Windows, or over an HTTP/TCP connection.
The package is the only one not using system calls to the `docker` CLI for managing images and containers and easily exposes connections to remote Docker daemons, which has to be configured on the Docker level otherwise.
Using the API gives access to more information and is system independent and likely more reliable than parsing command line output.
\CRANpkg{stevedore}'s own interface is automatically generated based on the OpenAPI specification of the Docker daemon, but still similar to the Docker CLI.
The interface is similar to R6 objects, in that a `stevedore_object` has a number of functions attached to it that can be called, and multiple specific versions of the Docker API can be supported thanks to the automatic generation\footnote{See \href{https://github.com/richfitz/stevedore/blob/master/development.md}{https://github.com/richfitz/stevedore/blob/master/development.md}.}.

**\CRANpkg{AzureContainers}** is an interface to a number of container-related services in Microsoft's [Azure Cloud](https://azure.microsoft.com/) \citep{AzureContainers_2019}.
While it is mainly intended for working with Azure, as a convenience feature it includes lightweight, cross-platform shells to Docker and Kubernetes (tools `kubectl` and `helm`).
These can be used to create and manage arbitrary Docker images and containers, as well as Kubernetes clusters on any platform or cloud service.

@@

## Deployment and continuous delivery
\label{deployment}

The cloud is the natural environment of containers, and subsequently containers are the go-to mechanism to deploy R server applications.
More and more continuous integration (CI) and continuous delivery (CD) services also use containers, opening up new options for use.
The controlled nature of containers, i.e., the possibility to abstract internal software environment from a minimal dependency outside of the container is crucial, for example to match test or build environments with production environments or transfer runnable entities to as-a-service infrastructures.

First, different packages use containers for the **deployment of R and \CRANpkg{Shiny} apps**.
\CRANpkg{Shiny} is a popular package for creating interactive online dashboards with R and it enables users with very diverse backgrounds to create stable and user friendly web applications.
For example, _ShinyProxy_ ([https://www.shinyproxy.io/](https://www.shinyproxy.io/)) is an open source tool to deploy Shiny apps in an enterprise context.
They feature single sign-on, but also in scientific use cases \citep[e.g., ][]{savini_epiexplorer_2019,glouzon_structurexplor_2017}.
ShinyProxy uses Docker containers to isolate user sessions and to achieve scalability for multi-user scenarious with multiple apps.
ShinyProxy itself is written in Java to accomodate corporate requirements and may itself run in a container for stability and availability.
The tool is build on ContainerProxy ([https://www.containerproxy.io/](https://www.containerproxy.io/)), which provides similar features for executign long-running R jobs or interactive R sessions.
The started containers can run on a regular Docker host but also in clusters.
Another example is the package \CRANpkg{golem} ([https://github.com/ThinkR-open/golem](https://github.com/ThinkR-open/golem)), which makes an heavy use of `dockerfiler` when it comes to creating the `Dockerfile` for building production-grade Shiny applications and deploying them.
\CRANpkg{googleComputeEngineR} ([https://cloudyr.github.io/googleComputeEngineR/](https://cloudyr.github.io/googleComputeEngineR/)) enables quick deployments of key R services, such as RStudio and Shiny, onto cloud virtual machines (VMs) with Google Cloud Compute Engine \citep{googleComputeEngineR_2019}.
The package utilises `Dockerfile`s to move the labour of setting up those services from the user to a premade Docker image, which is configured and run in the cloud VM.
For example, by specifying the template `template="rstudio"` in functions `gce_vm_template()/gce_vm()` an up to date RStudio Service image is launched for development work, while specifying `template="rstudio-gpu"` will launch an RStudio Server image with a GPU attached, etc.

@@

Third, containers can greatly facilitate the **deployment to existing infrastructures**.
_Kubernetes_ ([https://kubernetes.io/](https://kubernetes.io/)) is a container-orchestration system for managing container-based application deployment and scaling.
A _cluster_ of containers, orchestrated as a single deployment, e.g., with Kubernetes, can mitigate limitations on request volumes or a container occupied with computationally intensive task.
A cluster features load-balancing, autoscaling of containers across numerous servers (in the cloud or on premise), and restarting failed ones.
It may be your organisation has a Kubernetes cluster already for other applications, or you may use one of the different provides for managed Kubernetes clusters.
Docker containers are used within Kubernetes clusters to hold native code, for which Kubernetes creates a framework around network connections and scaling of resources up and down.
R applications, big parallel tasks, or scheduled batch jobs can be deployed in a scalable way using Kubernetes, and the deployment can even be triggered by changes to code repositories (i.e., CD), see blog post "R on Kubernetes" \citep{edmondson_r_2018}.
The package \pkg{googleKubernetesR} ([https://github.com/RhysJackson/googleKubernetesR](https://github.com/RhysJackson/googleKubernetesR)) is a proof of concept for wrapping the Google Kubernetes Engine API, Google's hosted Kubernetes solution, in an easy to use R package.
The package \CRANpkg{analogsea} provides a solution to programmatically create and destroy cloud VMs on the [Digital Ocean](https://www.digitalocean.com/) platform \citep{analogsea_2019}.
It also includes R wrapper functions to install Docker in such a VM, manage images, and control containers straight from R functions.
These functions are translated to Docker CLI commands and transferred transparently to the respective remove machine using SSH.
\CRANpkg{AzureContainers} is an umbrella package providing interfaces three commercial services of Microsoft's Azure Cloud, namely [Container Instances](https://azure.microsoft.com/en-us/services/container-instances/) for running individual containers, [Container Registry](https://azure.microsoft.com/en-us/services/container-registry/) for private image distribution, and [Kubernetes Service](https://azure.microsoft.com/en-us/services/kubernetes-service/) for orchestrated deployments.
While a package like \CRANpkg{plumber} provides the infrastructure for turning an R workflow a service, for production purposes it is usually necessary to take into account scalability, reliability and ease of management.
AzureContainers provides an R-based interface to these features and thereby simplifies complex infrastructure management to a number of R function calls, given an Azure account with sufficient credit\footnote{See \emph{"Deploying a prediction service with Plumber"} vignette for details:  \href{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01_plumber_deploy.html}{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01\_plumber\_deploy.html}.}.

The prevelance of Docker in industry naturally leads to usage of containers with R in such settings as well, as customers already manage platforms in Docker containers.
These products often entail a high amount of open source software in combination with proprietary layers adding the relevant commercializable features.
One such example is RStudio's data science platform [RStudio Team](https://rstudio.com/products/team/).
It allows teams of data scientists and their respective IT/DevOps groups to develop and deploy code in R and Python around the RStudio Open Source Server inside of Docker images, without requiring users to learn new tools or directly interact with containers.
The best practices for [running RStudio with Docker containers](https://support.rstudio.com/hc/en-us/articles/360021594513-Running-RStudio-with-Docker-containers) as well as [Docker images](https://github.com/rstudio/rstudio-docker-products) for RStudio's commerical products are publicly available.

@@

It deterministically builds a Docker image by parsing the contents of a repository, e.g., project dependency configurations or simple configuration files\footnote{See supported file types at \href{https://repo2docker.readthedocs.io/en/latest/config\_files.html}{https://repo2docker.readthedocs.io/en/latest/config\_files.html}.}.

@@

\citet{gentleman_statistical_2007} coined the term **Research Compendium** for a dynamic document together with supporting data and code.
They use the R package system \cite{core_writing_1999} as for the functional prototype to structuring, validation, and distribution of research compendia.
This concept has been taken up and extended\footnote{See full literature list at \href{https://research-compendium.science/}{https://research-compendium.science/}.}, not the least by applying containerisation and other methods for managing computing environments---see Section&nbsp;\nameref{envs}.
Containers give the researcher an isolated environment to assemble these research pipelines with specific versions of software.
Research workflows in containers are safe from contamination from other activities occuring on the researcher's computer, for example the installation of newest version of packages for teaching demonstrations.
Given the users in this scenario, i.e., often academics with limited formal software development training, templates and assistance with containers around research compendia is essential.
In many fields we see that a typical unit of research for a container is a research report or journal article, where the container holds the compendium, or self-contained set of data (or connections to data elsewhere) and code files needed to fully reproduce the article \citep{marwick_packaging_2018}.
The package \pkg{rrtools} ([https://github.com/benmarwick/rrtools](https://github.com/benmarwick/rrtools)) provides a template and convienence functions to apply good practices for research compendia, including a starter `Dockerfile`.
Images of compendium containers can be hosted on services such as Docker&nbsp;Hub for convienient sharing among collaborators and others. 
Similarly, packages such as \pkg{containerit} and \pkg{dockerfiler} can be used to manage the `Dockerfile` to be archived with a compendium on a data repository (e.g. [Zenodo](https://zenodo.org/), [Dataverse](https://dataverse.org/), [Figshare](https://figshare.com/), [OSF](https://osf.io/)).
A typial compendium's `Dockerfile` will pull a rocker image fixed to a specific version of R, and install R packages from the MRAN repository to ensure the package versions are tied to a specific date, rather than the most recent version.
A more extreme case is the _dynverse_ project ([https://dynverse.org/](https://dynverse.org/)), which packages over 50 computational methods with different environments (R, Python, C++, etc.) in Docker images, which can be executed from R.
_dynverse_ uses a CI (cf.&nbsp;\nameref{cicd}) to build Rocker-derived containers, test them, and push them to Docker&nbsp;Hub if the tests succeed.

Future researchers can download the compendium from the repository and run the included `Dockerfile` to build a new image that recreates the computational environment used to produce the original research results.
If building the image fails, the human-readable instructions in a `Dockerfile` are the starting point for rebuilding the environment.
Further safeguarding practices are currently under development or not part of common practice yet, such as preservation of images \citep{emsley_framework_2018} and storing them alongside `Dockerfile`s \citep[cf.][]{nust_opening_2017}, or pinning of system libraries beyond the tagged base images, which may be seen as stable or dynamic depending on the applied time scale \citep[see discussion on `debian:testing` base image in][]{RJ-2017-065}.
A recommendation of the recent National Academies' report on _Reproducibility and Replicability in Science_ is that journals _"consider ways to ensure computational reproducibility for publications that make claims based on computations"_ \citep{NASEM2019}.
In fields such as Political Science and Economics, journals are increasingly adopting policies requiring authors to publish the code and data required to reproduce computational findings reported in published manuscripts, subject to independent verification \citep{Jacoby2017,Vilhuber2019,Alvarez2018,Christian2018,Eubank2016,King1995}.
Problems with the computational environment, installation and availability of software dependencies are common.
R is gaining popularity in these communities, such as creating a research compendium.
In a sample of 105 replication packages published by the _American Journal of Political Science_ (AJPS) over 65% use R.
The NSF-funded Whole Tale project uses the Rocker Project community images with the goal of improving the reproducibility of published research artifacts and simplifying the publication and verification process for both authors and reviewers by reducing errors and time spent specifying the environment.

@@

Nonetheless, at least on the level of R packages some **consolidation** seems in order, e.g., to reduce the number of packages creating `Dockerfile`s from R code or controlling the Docker daemon with R code.
It remains to be seen which approach to control Docker, via the Docker API as \pkg{stevedore} or via system calls as \pkg{dockyard}/\pkg{docker}/\pkg{dockr}, is more sustainable, or if the question will be answered by the endurance of maintainers and sufficient funding.
Similarly, the capturing of environments and their serialization in form of a `Dockerfile` currently happens at different levels of abstraction and re-use of functionality seems reasonable, e.g., \pkg{liftr} could generate the environment with \pkg{containerit}, which in turn may use \pkg{dockerfiler} for low level R objects representing a `Dockerfile` and its instructions.
In this consolidation, the Rocker&nbsp;Project could play the role of coordinating entity.
Though for the moment, the sign of the times points to more experimentation and feature growth, e.g., images for GPU-based computing and artificial intelligence.
Even with coding being more and more accepted as a required, and achievable skill, an easier access, for example by exposing containerisation benefits via simple user interfaces in the users' IDE, could be an important next step.
Currently containerisation happens more in the background at the system level.
Furthermore, a maturing of the Rockerverse packages for managing containers may lead to their adoption where currently manual coding is required, e.g. in the case of \CRANpkg{RSelentium} or \CRANpkg{drake} (see Sections&nbsp;\nameref{rselenium} and \nameref{drake} respectively).
In specific cases, e.g., for \CRANpkg{analogsea}, the interaction with the Docker daemon may remain to specific to reuse first order packages to control Docker.

@@

The ordering of authors following DN and DE is alphabetical.
DN conceived the article idea, \href{https://github.com/nuest/rockerverse-paper/issues/3}{initialised the formation of the writing team}, wrote sections not mentioned below, and revised all sections.
DB wrote the section on \pkg{outsider}.
GD contributed the CARD.com use case.
DE wrote the introduction and the section about Containerization and Rocker and reviewed all sections.
RC \& EH contributed to the section on interfaces for Docker in R (_dynverse_ and `dynwrap`).
DC contributed content on Gigantum.
ME contributed to the section on processing and deployment to cloud services.
CF wrote paragraphs about \pkg{r-online}, \pkg{dockerfiler}, \pkg{r-ci} and \pkg{r-db}.
SL contributed content on RStudio's usage of Docker.
BM wrote the section on research compendia and made the project Binder-ready.
HN \& JN co-wrote the section on the T-Mobile use case.
KR wrote the section about \pkg{holepunch}.
NR wrote paragraphs about shared work environments and GPUs.
LS \& NT wrote the section on Bioconductor.
CW wrote the sections on Whole Tale and contributed the publication reproducibility audit use case.
NX contributed content on \pkg{liftr}.
All authors approved the final version.
This articles was collaboratively written at \href{https://github.com/nuest/rockerverse-paper/}{https://github.com/nuest/rockerverse-paper/}.
The \href{https://github.com/nuest/rockerverse-paper/graphs/contributors}{contributors page}, \href{https://github.com/nuest/rockerverse-paper/commits/master}{commit history}, and \href{https://github.com/nuest/rockerverse-paper/issues/}{discussion issues} provide a detailed view on the respective contributions.
